Formal Verification of Real-Time Data Processing of the LHC Beam Loss Monitoring System: A Case Study

### 2011-03-28 fmics/paper.tex
merge some dropbox changes

Figure~\ref{fig:theorems} gives a selection of theorems we have proved in HOL, some of which refer to additional constants defined in Figure~\ref{fig:auxdefs}.
Our central result, Theorem~(\ref{output_input_at_update_times}), equates the output of a slice to a sum of consecutive input counts.
More precisely, it says that if slice $n$ was just updated, then the output at tap $x$ is equal to the sum of $((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)$ input values, starting from the input $\mathsf{delay\_sum}\;n$ time steps ago.
In light of Theorem~(\ref{output_last_update}), which says a slice's output stays constant between update times for the slice, Theorem~(\ref{output_input_at_update_times}) characterises all outputs at all times.
Thus, the SRS structure's outputs are exact running sums of the input counts over windows whose sizes depend on the the width of the slice and position of taps.
However, the sums are delayed by the total delay across all previous slices (represented by $\mathsf{delay\_sum}$).

To prove Theorem~(\ref{output_input_at_update_times}), we first make the following observations\footnote{The precise statements of these observations are complicated by details of what happens when $t$ is low, or equivalently, when we are looking at shift register cells that have not yet received any input.}
\begin{enumerate}
\item
$\mathsf{SR}\;D\;n\;(m+1)\;t=\mathsf{SR}\;D\;n\;m\;(t-\mathsf{delay}\;n)$, since shift register cells get their values from the previous cell, but only update every $\mathsf{delay}\;n$ time steps;
\item
$\mathsf{SR}\;D\;n\;m\;t=\mathsf{SR}\;D\;n\;0\;(t-(m\times\mathsf{delay}\;n))$, by induction on $m$ using the result above;
\item
$\mathsf{output}\;D\;n\;x\;t$ is a sum of values of consecutive shift register cells (Theorem~(\ref{output_sum})), which can be proved by induction on $t$;
\item
Combining the last two results, $\mathsf{output}\;D\;n\;x\;t$ is a sum of consecutive values of the first shift register cell, $\mathsf{SR}\;D\;n\;0$;
\item
Thus, we can express $\mathsf{output}\;D\;n\;x\;t$ in terms of $\mathsf{output}\;D\;n'\;x'\;t$, where $(n',x')=\mathsf{input}\;n$, since the source of $\mathsf{SR}\;D\;n\;0$ is the previous (input) slice's output;
\item
Finally, by induction on $n$, we can express $\mathsf{output}\;D\;n\;x\;t$ as a function of $D$ alone, as in Theorem~(\ref{output_input_at_update_times}), using the result above for the inductive case.
\end{enumerate}

>>>>

To prove Theorem~(\ref{output_input_at_update_times}), we first make the following observations\footnote{The precise statements of these observations are complicated by details of what happens when $t$ is low, or equivalently, when we are looking at shift register cells that have not yet received any input.}
%Figure~\ref{fig:theorems} gives a selection of theorems we have proved in HOL, some of which refer to additional constants defined in Figure~\ref{fig:auxdefs}.
%Our central result, Theorem~(\ref{output_input_at_update_times}), equates the output of a slice to a sum of consecutive input counts.
Our central result equates the output of a slice to a sum of consecutive input counts.
More precisely, it says that if slice $n$ was just updated, then the output at tap $x$ is equal to the sum of $((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)$ input values, starting from the input $\mathsf{delay\_sum}\;n$ time steps ago.
\begin{theorem}

The output of slice n at update time t ... (statement of the theorem)
\begin{multline}\label{output_input_at_update_times}
\forall{D\,n\,x\,t}.\,{0 < n}\land{\mathsf{update\_time\;n\;t}}\implies\\\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)-1}\begin{cases}0&t<m+\mathsf{delay\_sum}\;n\\D\;(t-m-\mathsf{delay\_sum}\;n)&\text{otherwise}\end{cases}
\end{multline}
\end{theorem}

\begin{proof}
\begin{enumerate}
\item
$\mathsf{SR}\;D\;n\;(m+1)\;t=\mathsf{SR}\;D\;n\;m\;(t-\mathsf{delay}\;n)$, since shift register cells get their values from the previous cell, but only update every $\mathsf{delay}\;n$ time steps;
\item
$\mathsf{SR}\;D\;n\;m\;t=\mathsf{SR}\;D\;n\;0\;(t-(m\times\mathsf{delay}\;n))$, by induction on $m$ using the result above;
\item
$\mathsf{output}\;D\;n\;x\;t$ is a sum of values of consecutive shift register cells (Theorem~(\ref{output_sum})), which can be proved by induction on $t$;
\item
Combining the last two results, $\mathsf{output}\;D\;n\;x\;t$ is a sum of consecutive values of the first shift register cell, $\mathsf{SR}\;D\;n\;0$;
\item
Thus, we can express $\mathsf{output}\;D\;n\;x\;t$ in terms of $\mathsf{output}\;D\;n'\;x'\;t$, where $(n',x')=\mathsf{input}\;n$, since the source of $\mathsf{SR}\;D\;n\;0$ is the previous (input) slice's output;
\item
Finally, by induction on $n$, we can express $\mathsf{output}\;D\;n\;x\;t$ as a function of $D$ alone, as in Theorem~(\ref{output_input_at_update_times}), using the result above for the inductive case.
\end{enumerate}

\end{proof}

In light of Theorem~(\ref{output_last_update}), which says a slice's output stays constant between update times for the slice, Theorem~(\ref{output_input_at_update_times}) characterises all outputs at all times.
Thus, the SRS structure's outputs are exact running sums of the input counts over windows whose sizes depend on the the width of the slice and position of taps.
However, the sums are delayed by the total delay across all previous slices (represented by $\mathsf{delay\_sum}$).
